### 2018-03-11 paper/3-compute.tex
Add some pieces for Andrey to include
Work on S3

\begin{minted}[xleftmargin=10pt]{haskell}
type Compute c k v = @\std{forall}@ f. c f => (k -> f v) -> k -> Maybe (f v)
type Build c i k v = Compute c k v -> k -> Maybe i -> (k -> v) -> (i, k -> v)
\end{minted}

\noindent
Compute represents build rules (in build systems) or formulas (in spreadsheets),
and is completely isolated from the world of compilers, file systems, memories,
caches, and all other complexities of real build systems. We explain the compute
type in~\S\ref{sec-general-compute}, and then discuss several noteworthy special cases:
functorial (\S\ref{sec-compute-functor}), applicative
(\S\ref{sec-compute-applicative}) and monadic (\S\ref{sec-compute-monad}) types
of compute.

Build corresponds to the algorithm used for bringing a key-value store to a
consistent state by recomputing dirty values in an appropriate order. We explain
the build type in~\S\ref{sec-build}, and then provide implementations for several
build systems in~\S\ref{sec-implementations}.

@@

In this section we introduce a common vocabulary for build systems that allows
us to abstract away from specific application domains, such as software
development or spreadsheets. We will use this vocabulary throughout the rest of
the paper.

\emph{Keys} are used to distinguish \emph{values}. In software build systems
keys are typically filenames, e.g. \cmd{main.c}, whereas values are file
contents (a C program source code in this case). In spreadsheets keys are cell
names, e.g. \cmd{A1}, and values are numbers, text, etc. that are typically
displayed inside cells. In Haskell code, we will use type variables \hs{k}
and \hs{v} to denote keys and values, respectively.

We use a cryptographic \emph{hash function} \hs{hash :: v -> Hash} for efficient
tracking and sharing of build results, where \hs{Hash} is an abstract data type
equipped with the equality test.

A \emph{store} associates keys to values. It is convenient to assume that a store
is total, i.e. it contains a value for every possible key. We therefore also
assume that the type of values is capable of encoding values corresponding to
non-existent keys (missing files or empty cells). This, in particular, suggests
that it is possible to \emph{depend on the absence of a value} which is a useful
feature for build systems. In addition to usual \hs{get} and \hs{put} operations,
some stores support the \hs{getHash} operation implemented more efficiently than
by hashing the result of \hs{get}. For example, GVFS (Git Virtual File
System)~\cite{gvfs} stores file hashes and downloads file contents only on demand.

>>>>

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
type Compute c k v = @\std{forall}@ f. c f => (k -> f v) -> k -> Maybe (f v)
type Build c i k v = Compute c k v -> k -> Maybe i -> (k -> v) -> (i, k -> v)
\end{minted}
\vspace{1mm}

\noindent
\hs{Compute} represents an abstract set of \emph{build tasks}, for example, build
rules in software build systems or formulas in spreadsheets. It is completely
isolated from the world of compilers, file systems, calculation chains, memories,
caches, and all other complexities of real build systems. We explain the compute
type in~\S\ref{sec-general-compute}, and then discuss several noteworthy special
cases: functorial (\S\ref{sec-compute-functor}), applicative
(\S\ref{sec-compute-applicative}) and monadic (\S\ref{sec-compute-monad}) types
of compute.

\hs{Build} corresponds to the algorithm used for bringing a key-value store
\store to a consistent state by recomputing out-of-date values in an
appropriate order. We explain the build type and use it to develop a formal
definition of build systems in~\S\ref{sec-general-build}.
Sections~\S\ref{sec-build} and~\S\ref{sec-implementations} further deconstruct
build algorithms and provide concrete implementations for several build systems.

@@

We start by introducing a common vocabulary for build systems that allows
us to abstract away from specific application domains, such as software
development or spreadsheets. We will use this vocabulary throughout the rest of
the paper.

\emph{Keys} are used to distinguish \emph{values}. In software build systems,
keys are typically filenames, e.g.~\cmd{main.c}, whereas values are file
contents (a C program source code in this case). In spreadsheets, keys are cell
names, e.g. \cmd{A1}, and values are numbers, text, etc. that are typically
displayed inside cells. In Haskell code, we will use type variables \hs{k}
and \hs{v} to denote keys and values, respectively.

We use a cryptographic \emph{hash function} \hs{hash :: v -> Hash} for efficient
tracking and sharing of~build results, where \hs{Hash} is an abstract data type
equipped with the equality test.

A \emph{store} associates keys to values. It is convenient to assume that a
store is total function \store, i.e. it contains a value for every possible key.
We therefore also assume that the type of values is capable of encoding values
corresponding to non-existent keys (missing files or empty cells). This, in
particular, suggests that it is possible to \emph{depend on the absence of a
value}, which is a useful feature for build systems. In addition to usual
\hs{get} and \hs{put} operations, some stores support the \hs{getHash} operation
implemented more efficiently than by hashing the result of \hs{get}. For example,
GVFS (Git Virtual File System)~\cite{gvfs} stores file hashes and downloads file
contents only on demand -- an important feature for cloud build systems.
